---
title: "Untitled"
author: "Van Gonzales"
date: "10/29/2021"
output: html_document
---
## Overview
Using this data we will compare the mean sizes of purple urchins in the control treatments and annual treatments within sections of the Naples Reef in 2010.

### Data Citation
 Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 
 
### Setup
```{r setup, include=FALSE, message = FALSE, warning = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #attaching packages
library(here)
library(janitor)
library(kableExtra)
library(lubridate)

```

### Reading in & Cleaning up the Data
```{r}
urchins <- read_csv(here("data", "urchins.csv")) #read in data

urchins_clean <- urchins %>% #putting column names into lower case
  clean_names()
```

### Wrangling Data
Only including purple urchins from the Naples Reef observed in 2010
```{r}
purple_naples <- urchins_clean %>% 
  filter(site == "NAPL", 
         common_name == "Purple Urchin") %>% #filter for purple & NAPL
  mutate(date = mdy(date)) %>% #use mdy() to convert to date class
  mutate(year = year(date)) %>% 
  filter(year == "2010") %>% #filter for only 2010 data
  mutate(treatment = case_when(
    treatment == "ANNUAL" ~ "Annual",
    treatment == "CONTROL" ~ "Control")) #do this for easier organization
```

### Data Visualization
Creating a histogram and a qq plot by treatment, to display purple urchins in 2010.
```{r}
ggplot(data = purple_naples, aes(x = size)) + #creating ggplot
  geom_histogram(bins = 10, color = "purple",
                 aes(fill = treatment)) +
                   facet_wrap(~ treatment) +
                   labs(x = "Size (cm)", y = "Count", 
                    title = "Purple Sea Urchin Size (cm) by Treatment",
                    fill = "Treatment Type") #all customizations

ggplot(data = purple_naples, aes(sample = size)) +
  stat_qq() + #creating qq plot
  geom_qq_line() +
  facet_wrap(~treatment)
```
Data Analysis: Based on the graphs, the distributions look very similar. There doesn't seem to be a huge difference between the two treatment types.

```{r}
urchin_treatment <- purple_naples %>% 
  group_by(treatment) %>%  #grouping for summary table
  summarize(mean_size = round(mean(size), 2), median_size = median(size),
            sd_size = round(sd(size), 2),
            sample_size = n())

urchin_treatment %>% 
  kable(col.names = c("Treatment", #creating column names summary table
                      "Mean Size (cm)",
                      "Median Size (cm)",
                      "Standard Deviation (cm)",
                      "Sample Size (n)"),
        caption = "Figure 1. Purple sea urchin size (cm) by treatment type on Naples reef in 2010.") %>% 
 kable_classic(full_width = F, html_font = "Cambria") #just styling
```













